---
title: "Survival Example"
author: "Eric Dunipace"
date: "6/9/2019"
output: html_document
---
<style>
p.caption {
  font-size: 0.9em;
  font-style: italic;
  color: grey;
  margin-right: 10%;
  margin-left: 10%;  
  text-align: justify;
}
</style>
```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(SparsePosterior)
require(CoarsePosteriorSummary)
require(ggplot2)
require(ggsci)
require(ggridges)
require(doParallel)
require(xtable)
require(rjags)
```
<!-- start running of code -->
```{r data, message= FALSE, warning = FALSE, echo=FALSE}
#### Load packages ####
require(SparsePosterior)
require(CoarsePosteriorSummary)
require(ggplot2)
require(ggsci)
require(doParallel)
require(xtable)
require(rjags)
require(ggridges)
require(rstan)

#### generate data ####
set.seed(7848)
n <- 2^10 # 1024
p <- 101 # must be at least 6
nsamp <- 1000
nthin <- 20
nlambda <- 100
lambda.min.ratio <- 1e-3
gamma <- 1
pseudo.observations <- 0

#retrieve parameters
target <- get_survival_linear_model()
target$X$corr <- 0.5

# generate matrices
X <- target$X$rX(n, target$X$corr, p)
param <- target$rparam() #for non-linear model, not needed
p_star <- length(param)
listdat <- target$rdata(n, X[,1:p_star], param, corr.x = target$X$corr, scale = TRUE)
Y <- listdat$Y
true.log.rate <- listdat$eta
true.rate <- listdat$mu

```
```{r posterior estimation, message=FALSE, results='hide', warning = FALSE, echo = FALSE}
hyperparameters <- list(mu = 0.0, sigma = 1,
                        prior_frac = 0.1,
                        spike_a = 1 ,
                        spike_b = 1)
jags_dir <- "../../exec/Jags/survival.jags"

stan_dir <- "../../exec/Stan/cox_horseshoe.stan"

postFile <- "surv_post.rds"
if(!file.exists(postFile)) {
  # post.surv <- target$rpost(nsamp, X,Y, hyperparameters = hyperparameters, 
  #                           jags_dir = jags_dir, thin = nthin)
  post.surv <- target$rpost(nsamp, X,Y, hyperparameters = hyperparameters, 
                            stan_dir = stan_dir, thin = nthin, chains = 1,
                            m0 = 0.1, method = "stan")
  saveRDS(post.surv, file = postFile)
} else {
  post.surv <- readRDS(postFile)
}

```
```{r coarsened posterior, echo = FALSE, message=FALSE, warning= FALSE}
penfact <- rep(1, p)

adaptFile <- "surv_adapt.rds"
if(!file.exists(adaptFile)){
  selection <- W2L1(X = X, Y = NULL, #post.surv$eta, 
                        theta=post.surv$theta, penalty="selection.lasso",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1e4, maxit = 5e2, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="selection.variable")
  projection.lasso <- W2L1(X = X, Y = post.surv$eta, 
                        theta=post.surv$theta, penalty="lasso",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1, maxit = 5e2, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="projection")
  projection.mcp <- W2L1(X = X, Y = post.surv$eta, 
                        theta=post.surv$theta, penalty="mcp",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1, maxit = 1e3, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="projection")
  saveAdapt <- list(selection = selection, lasso = projection.lasso,
                    mcp = projection.mcp)  
  saveRDS(saveAdapt, file = adaptFile)
} else {
  saveAdapt <- readRDS(adaptFile)
  
  selection <- saveAdapt$selection
  projection.lasso <- saveAdapt$lasso
  projection.mcp <- saveAdapt$mcp
  
}

```
```{r w2calc, echo=FALSE, message=FALSE, warning= FALSE}
compare <- list("Selection" = selection,
                "Projection Lasso" = projection.lasso,
                "Projection MCP" = projection.mcp)
w2file <- "w2plot_surv.rds"
if(!file.exists(w2file)) {
  w2plot <- plot.compare(compare, post.surv$eta, X, t(post.surv$theta), 
                          "w2", "mean", TRUE)
  w2plotData <- w2plot$data$mean
  saveRDS( w2plotData, w2file )
  w2plots <- w2plot$plot$mean
} else {
  w2plotData <- readRDS(w2file)
  w2plotData$groups <- factor(w2plotData$groups)
  w2plots <- ggplot(w2plotData, 
                    aes(x=nactive, y=dist, color = groups, group=groups )) +
        geom_line() + scale_color_jama() + labs(color ="Method") +
        xlab("Number of active coefficients") + ylab("2-Wasserstein Distance") + theme_bw() +
        scale_x_continuous(expand = c(0, 0)) +
        scale_y_continuous(expand = c(0, 0),
                           limits = c(0, max(w2plotData$dist)*1.1))
}



```
```{r msecalc, echo = FALSE, warning= FALSE, message=FALSE}
compare <- list("Selection" = selection,
                "Projection Lasso" = projection.lasso,
                "Projection MCP" = projection.mcp)
msefile <- "mseplot_surv.rds"
if(!file.exists(msefile)) {
  mseplot <- plot.compare(compare, true.rate, X, t(post.surv$theta), 
                          method = "mse", quantity = "mean", 
                          parallel=TRUE, transform = exp)
  mseplotData <- mseplot$data$mean
  saveRDS( mseplotData, msefile )
  mseplots <- mseplot$plot$mean
} else {
  mseplotData <- readRDS(msefile)
  mseplotData$groups <- factor(mseplotData$groups)
  mseplots <- ggplot(mseplotData, 
                    aes(x=nactive, y=dist, color = groups, group=groups )) +
        geom_line() + scale_color_jama() + labs(color ="Method") +
        xlab("Number of active coefficients") + ylab("2-Wasserstein Distance") + theme_bw() +
        scale_x_continuous(expand = c(0, 0)) +
        scale_y_continuous(expand = c(0, 0),
                           limits = c(0, max(mseplotData$dist)*1.1))
}

```
```{r calc_ridge, message= FALSE, warning = FALSE, echo=FALSE}
compare <- list("Selection" = selection,
                "Projection Lasso" = projection.lasso,
                "Projection MCP" = projection.mcp)

rank.file <- "ranks_surv.rds"
if(!file.exists(rank.file)){
  quantRanks <- ranking(compare, full = post.surv$eta, p=2, maxCoef = 5,
                      quantiles = c(0, 0.25, 0.5, 0.75, 1), 
                      transform = exp)
  saveRDS(quantRanks, file=rank.file)
} else {
  quantRanks <- readRDS(rank.file)
}

ridgePlots <- ridgePlot(compare, 
                       maxCoef = quantRanks$ncoef,
                       index = quantRanks$index,
                       scale = 2, alpha = 0.5,
                       full = post.surv$eta, 
                       transform = exp)
```
<!-- end running of code -->
Our next example is in a setting with Exponentially distributed data. We generate $n = `r n`$ predictor variables, $X$, from a Multivariate Normal,
\[ X_i \sim N(0, \Sigma), \] with $\Sigma_{i,i} = 1$ for all $i$ from 1 to $p = `r p-1`.$ The first 5 dimensions of $X$ are correlated, dimensions 6-10 are correlated, dimensions 10-20 are correlated, dimensions 21-50 are correlated, and dimensions 50-p = `r p-1` are correlated with correlations `r target$X$corr`. Thus, $\Sigma$ is block diagonal.

### Parameters
We generate parameters as follows
\[\beta_{0} \sim \text{Unif}(1,2), \]
\[ \beta_{1:5} \sim \text{Unif}(0.1,0.2),\]
\[ \beta_{6:10} \sim \text{Unif}(-0.2,-0.1), \] and
\[ \beta_{11:20} \sim \text{Unif}(0,0.05), \] 
so that only the first 20 dimensions of $X$ are relevant to predicting $Y$

### Outcome
We sample the outcome as
\[Y_i \sim \text{Expo}(\lambda_i), \]
with $\lambda_i = \beta_0 + X_{i, 1:20} \beta_{1:20}$.

```{r data_print, eval=FALSE}
<<data>>
```
## Posterior Estimation
We put conjugate priors on our parameters,
\[
  \beta_{j} \sim \alpha_j N(0,1) + (1-\alpha_j) \delta_0
\] and
\[
  \alpha_j \sim \text{Beta}(1,1),
\]
for all $j$ from $1,...,p$. The intercept has a simple, non-informative prior
\[ \beta_{0} \sim \alpha_j N(0,10^2).\]
Then we estimate the posterior using JAGS (Just Another Gibbs Sampler). The following code sets the hyperparameters and runs rjags:
```{r post print, eval=FALSE}
hyperparameters <- list(mu = 0.0, sigma = 1,
                        prior_frac = 0.1,
                        spike_a = 1 ,
                        spike_b = 1)
jags_dir <- "../../exec/Jags/survival.jags"

post.surv <- target$rpost(nsamp, X,Y, hyperparameters = hyperparameters, 
                            jags_dir = jags_dir, thin = nthin)
```
From the posterior model, only two coefficients are reliably selected---$\beta_{ `r which(apply(post.surv$alpha,2,median)==1)[1] `}$ and $\beta_{`r which(apply(post.surv$alpha,2,median)==1)[2] `}$. Further only `r sum(colMeans(post.surv$alpha) !=0)` coefficients are selected into the model at all. Thus, we won't have to work with all `r p` coefficients.

## Coarsened Approximations to the Full Posterior
Then we calculate the coarsened versions of our posterior using 2 methods

1. The selection method
2. The projection method with a lasso penalty and a minimax-concave penalty (MCP)

```{r coarse calc, eval=FALSE}
selection <- W2L1(X = X, Y = NULL, #post.surv$eta, 
                        theta=post.surv$theta, penalty="selection.lasso",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1e4, maxit = 5e2, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="selection.variable")
projection.lasso <- W2L1(X = X, Y = post.surv$eta, 
                        theta=post.surv$theta, penalty="lasso",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1, maxit = 5e2, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="projection")
projection.mcp <- W2L1(X = X, Y = post.surv$eta, 
                        theta=post.surv$theta, penalty="mcp",
                        nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
                        infimum.maxit=1, maxit = 1e3, gamma = gamma,
                        pseudo_observations = pseudo.observations, 
                        display.progress = TRUE,
                        penalty.factor = penfact, method="projection")
```


## Distance Plots
Next we calculate and plot two things: 1) the 2-Wasserstein distance between each coarsened posterior and 2) the full posterior and the MSE between the true rate parameter and the version calculated from the coarsened posteriors. We generate the data for the plots and plots themselves with the following code:

```{r plot run, eval = FALSE}
compare <- list("Selection" = selection,
                "Projection Lasso" = projection.lasso,
                "Projection MCP" = projection.mcp)

w2plot <- plot.compare(models = compare, target = post.surv$eta, 
                       method = "w2", quantity = "mean",
                          parallel=TRUE)
mseplot <- plot.compare(models = compare, target = true.rate,
                          method = "mse", quantity = "mean", 
                          parallel=TRUE, transform = exp)

```

Which gives us these plots
```{r plots, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 4, fig.width=4, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='2-Wasserstein Distance and MSE to the true rate parameters for each method.'}
print(w2plots)
print(mseplots)
```

We can see that as more coefficients are added, the 2-Wasserstein distance between the coarsened model and the full models decrease. Similarly, the MSE decreases sharply for the first few coefficients but then only slowly decreases afterward. This may suggest the intercept is the most important variable, which for this data, it is.

## Ridgline Plots and Visualizing Observation-level Predictive Distributions

Finally, we may want to know how well or how band our predictions are doing for individual posterior predictive distributions, so we can look at ridgeline plots for individuals for our desired covariate size. Further, we can see the best performing coarsened posterior predictive and the worst performing posterior predictive, and anywhere inbetween!

Let's say we've decided 5 coefficients is all we want to communicate or make predictions with. How good or how bad will our predictive distribution be? This code will calculate the ranks
```{r rank print, eval=FALSE}
quantRanks <- ranking(compare, full = post.surv$eta, p=2, maxCoef = 5,
                      quantiles = c(0, 0.25, 0.5, 0.75, 1),
                      transform = exp)
```
Note that we transform the distributions to be on the scale of the rate parameter rather than the log-rate parameter, but we don't have to do this. We've chosen to look at the individuals with the best and worst performing posterior predictive linear predictors, as well as those individuals at the $25^{th},\, 50^{th}, \text{ and } 75^{th}$ percentiles. Once we have our ranks, we can make ridgeline plots of the distributions with this code
```{r ridge print, eval=FALSE, mesage=FALSE, warning=FALSE}
ridgePlots <- ridgePlot(compare, 
                       maxCoef = quantRanks$ncoef,
                       index = quantRanks$index,
                       scale = 2, alpha = 0.5,
                       full = post.surv$eta, 
                       transform = exp)
```

Then we can look at the ridgeline plots for each returned individual to see the discrepancy for 5 coefficients:
```{r ridge plots 1, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 6, fig.width=8, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='Ridgeline plots of the posterior predictive distribution for the observation with the smallest 2-Wasserstein distance to the distribution of the rate parameter for full model.'}
print(ridgePlots[[1]])
```
```{r ridge plots 2, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 6, fig.width=8, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='Ridgeline plots of the posterior predictive distribution for the observation at the $25^{th}$ percentile of 2-Wasserstein distances to the distribution on the rate parameter for the full model.'}
print(ridgePlots[[2]])
```
```{r ridge plots 3, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 6, fig.width=6, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='Ridgeline plots of the posterior predictive distribution for the observation at the $50^{th}$ percentile of 2-Wasserstein distances to the distribuion of the rate parameter for the full model.'}
print(ridgePlots[[3]])
```
```{r ridge plots 4, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 6, fig.width=8, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='Ridgeline plots of the posterior predictive distribution for the observation at the $75^{th}$ percentile of 2-Wasserstein distances to the distribution of the rate parameter for the full model.'}
print(ridgePlots[[4]])
```
```{r ridge plots 5, echo=FALSE, message=FALSE, warning=FALSE, fig.height= 6, fig.width=8, out.width="49%", optipng = '-o7', fig.align = "center", fig.show='hold', fig.cap='Ridgeline plots of the posterior predictive distribution for the observation with the largest 2-Wasserstein distance to the distribution of the rate parameter for the full model.'}
print(ridgePlots[[5]])
```
